---
title: "16S Demultiplexing"
output: html_document
date: "2023-06-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tools)
library(fs)
library(here)
library(Biostrings)
```


```{r}
"config-mimic.R" %>%
    here() %>%
    source()

# make directory for output
if (dir_exists(demux_dir)) {dir_delete(demux_dir)}
dir_create(demux_dir)

```

## Check Data Integrity

```{r md5sums check}

mimic_md5 <- mimic_md5sum %>%
  read_delim(delim=" ", 
             trim_ws = TRUE,
             col_names = c("md5sum", "X2" ,"filename")) %>%
  select(-X2)


md5check <- mimic_data_dir %>%
  list.files(full.names = TRUE) %>%
  md5sum %>%
  enframe %>%
  dplyr::rename(fullpath=name, observed_md5=value) %>%
  mutate(filename=basename(fullpath)) %>%
  right_join(mimic_md5, by="filename") %>%
  mutate(match=identical(md5sum,observed_md5))
    
md5check %>%
    pull(match) %>%
    all ->
    all_md5s_ok

stopifnot(all_md5s_ok)

```

## Instructions for downloading ea-utils for fastq-multx
1. In the terminal, run "module load miniconda3"
2. In the terminal, run "conda create --name demux ea-utils" (demux refers to the name of the new environment and ea-utils is the package installed in it)
3. In the markdown file, create a new code block, but instead of {r}, use {bash engine.opts='-l'} and in the first line run "conda activate demux", this loads in the demux environment and lets you use ea-utils in that code block


## Demultiplexing

```{bash engine.opts='-l'}

conda activate demux

fastq-multx

```


```{r barcode table}

mimic.metadata <- read.table(map_file) %>%
  select(`Sample` = `V1`,
         BarcodeSequence = `V2`)  %>%
  write_delim(barcode_table,        # output barcodes to a file 
              delim="\t", 
              col_names=FALSE)

```

```{bash engine.opts='-l'}

conda activate demux


set -u
fastq-multx \
  -m 3 \
  -d 2 \
  -x \
  -B $BARCODE_TABLE \
  $BARCODE_FASTQ \
  $R1_FASTQ \
  $R2_FASTQ \
  -o ${DEMUX_DIR}/'%_I1.fastq.gz' \
  -o ${DEMUX_DIR}/'%.R1.fastq.gz' \
  -o ${DEMUX_DIR}/'%.R2.fastq.gz' > \
  $DEMUX_STDOUT

```

```{r}
demux_stdout %>%
  read_tsv(file=.,
           skip=1, 
           col_names=c("Id", "Count", "I1_path", "R1_path", "R2_path")) ->
  reads_per_sample

```

```{r}
reads_per_sample %>%
    mutate(groupval = 
           ifelse(str_detect(Id,"unmatched|total"),
                 Id,
                 "sample")) %>%
    group_by(groupval) %>%
    summarise(Count=sum(Count))

```

## Rerun with reverse complement barcodes

```{r}

if (dir_exists(demux_dir)) {dir_delete(demux_dir)}
dir_create(demux_dir)

```

```{r}

read_table(map_file) %>% # read in map file
  select(Sample = `#SampleID`, 
         BarcodeSequence = `BarcodeSequence`) %>%   
  deframe %>%                          # convert to a named vector (expected input for DNAStringSet constructor)
  DNAStringSet %>%                     # convert to DNAStringSet
  reverseComplement %>%                # reverse complement the barcodes
  as.data.frame %>%                    # convert to dataframe for write_delim
  rownames_to_column %>% 
  write_delim(rc_barcode_table,        # output barcodes to a file 
              delim="\t", 
              col_names=FALSE)


```


```{bash engine.opts='-l'}

conda activate demux


set -u
fastq-multx \
  -m 3 \
  -d 2 \
  -x \
  -B $RC_BARCODE_TABLE \
  $BARCODE_FASTQ \
  $R1_FASTQ \
  $R2_FASTQ \
  -o ${DEMUX_DIR}/'%_I1.fastq.gz' \
  -o ${DEMUX_DIR}/'%.R1.fastq.gz' \
  -o ${DEMUX_DIR}/'%.R2.fastq.gz' > \
  $RC_DEMUX_STDOUT

```

```{r}
rc_demux_stdout %>%
  read_tsv(file=.,
           skip=1, 
           col_names=c("Id", "Count", "I1_path", "R1_path", "R2_path")) ->
    reads_per_sample_rc_barcode

```


```{r}
reads_per_sample_rc_barcode %>%
    mutate(groupval = 
           ifelse(str_detect(Id,"unmatched|total"),
                 Id,
                 "sample")) %>%
    group_by(groupval) %>%
    summarise(Count=sum(Count))

```

# Rerunning with reverse complement barcodes increased the unmatched substantially, stick with the original demux

