---
title: "Power calculation for the effect of abx and saline gavage on PD1 vs IgG on mc38 tumor volume"
author: "Dan Spakowicz"
date: '2022-10-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(effsize)
library(pwr)
```

# Summary

This script reads in tumor volume data from a saline gavage experiment, selects a time point where all mice were alive (15 days after mc38 injection) and calculates the Cohen's d effect size for the difference between PD1 and IgG. It then uses the effect size to calculate the power of detecting a difference in tumor volume for different group sizes.

# Analysis

Read in tumor volume data from previous experiments

```{r, message=FALSE, warning=FALSE}
x <- read_csv("../data/2022-08-11_tumor-size_R-formatted.csv")

head(x)
```

```{r}
saline <- 
  x %>%
  filter(grepl("Saline", microbiome.sample))
```

Visualize the saline data

```{r}
saline %>%
  ggplot(aes(x = days.from.injection, y = tumor.volume)) +
    geom_line(aes(group = unique.mouse.id, color = treatment)) + 
  theme_bw() +
  geom_smooth(aes(color = treatment))
```

```{r}
day15 <- 
  saline %>%
  filter(days.from.injection == 15)

day15 %>%
  group_by(treatment) %>%
  summarize(mean = mean(tumor.volume),
            sd = sd(tumor.volume))
```

Calculate the effect size using Cohen's *d*

```{r}
cohen.d(day15$tumor.volume, day15$treatment)

cohd <- cohen.d(day15$tumor.volume, day15$treatment)$estimate
```

Calculate the power with 6 mice per group

```{r}
pwr.t.test(n = 6, d = cohd, sig.level = 0.05, alternative = "less")
```

Check a range of sample sizes

```{r}
# Wrap into function
mousePower <- function(n) {
  pwr.t.test(n = n, d = cohd, sig.level = 0.05, alternative = "less")$power
}

# Calculate power over a range of sample sizes
n.range <- seq(1, 24)
power <- lapply(n.range, mousePower) %>%
  unlist

df <- 
  data.frame(n = n.range,
             power = power)

```

```{r, fig.cap="Figure 1. The power to detect a difference between groups at a range of sample sizes."}
df %>%
  ggplot(aes(n, power)) +
  geom_line(lwd = 1, aes(color = power), show.legend = FALSE) +
  theme_bw() +
  labs(x = "Number of Mice per Group",
       y = "Power") +
  scale_color_viridis_c(direction = -1)
```
