---
title: "Mimic24 Tumor Volume and Survival"
author: "Dan Spakowicz"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(survminer)
library(survival)
library(ggsurvfit)

# Set paths to locate OneDrive shared mimic folder
source("00-paths.R")
```

```{r}
# x <- readxl::read_xlsx("../data/tumor-size/AOT00055226-mimic24/mimic24_tumor-size.xlsx")
x <- readxl::read_xlsx(file.path(paths$onedrive, 
                                 "data", 
                                 "AOT00055226-mimic24", 
                                 "mimic24_tumor-size.xlsx"))

head(x)
```

```{r}
y <- 
  x %>%
  drop_na(microbiome.sample) %>%
  mutate(tumor.volume = (`tumor.length.(mm)` * (`tumor.width.(mm)`^2)) / 2) %>%
  mutate(mouse = paste(tail.number, cage, sep = "."))

head(y)  
```

```{r, warning=FALSE}
y %>%
  ggplot(aes(x = date, y = tumor.volume)) +
  geom_line(aes(group = mouse,
                color = microbiome.sample)) +
  theme_bw() +
  facet_wrap(~microbiome.sample)

```

```{r}
y %>%
  ggplot(aes(x = date, y = tumor.volume)) +
  geom_smooth(aes(color = microbiome.sample)) +
  theme_bw() +
  facet_wrap(~microbiome.sample)

```

```{r}

y %>%
  ggplot(aes(x = date, y = tumor.volume)) +
  geom_smooth(aes(color = microbiome.sample)) +
  theme_bw() 

```

# Longitudinal modeling

Check the distribution of the response variable

```{r}
y %>%
  ggplot(aes(x = tumor.volume)) +
  geom_histogram()

y %>%
  ggplot(aes(x = log(tumor.volume))) +
  geom_histogram()
```

A log transformation looks good.

Now I'll create the time variables

```{r}
z <- 
  y %>%
  mutate(microbe = fct_relevel(microbiome.sample, "PBS")) %>%
  mutate(log.tumor.volume = log(tumor.volume+1)) %>%
  mutate(days = as.Date(date) - as.Date("2024-03-14", "%F")) %>%
  mutate(days = as.numeric(days)) %>%
  mutate(days2 = days^2)
```

```{r}
m1 <- lm(log.tumor.volume ~ days + days2 * microbe,
         data = z)

summary(m1)
```




```

```{r}
surv.in <- 
  x %>%
  mutate(mouse = paste(tail.number, cage, sep = ".")) %>%
  # this should work but also removes NA... 
  # == "n/a" only selects those 6 though
  # max days on trial should be 24 
  filter(`tumor.length.(mm)` != "n/a") %>%
    group_by(mouse) %>%
  mutate(tumor.length = as.numeric(`tumor.length.(mm)`),
         tumor.width = as.numeric(`tumor.width.(mm)`)) %>%
  mutate(days.on.trial = max(date) - min(date)) %>%
  ungroup() %>%
  filter(!is.na(microbiome.sample)) %>%
  select(mouse, days.on.trial, microbiome.sample, tumor.length) %>%
  distinct() %>%
  mutate(status = ifelse(days.on.trial < max(days.on.trial), 1, 0)) %>%
  mutate(status = if_else(tumor.length > 16,
                          true = 1,
                          false = status))

head(surv.in)
```

```{r}
surv.m <- ggsurvfit::survfit2(Surv(days.on.trial, status) ~ microbiome.sample, 
                      data = surv.in)

surv.m %>%
  ggsurvfit() +
  add_pvalue()

ggsave("../figures/mimic24_survival.png",
       height = 4, width = 6)
```
